Comparing Exchangeable Effects (EE) and Exchangeable Standardized Effects (ET) models

### 2016-10-19 analysis/CompareEEvsETmodel.Rmd
Completed revisions of CompareEEvsETmodel.Rmd; it is working with the new version of ashr.

```{r generate_data}
set.seed(1234)
nsamp      = 1000
betahat.se = rgamma(nsamp,1,1)
beta       = betahat.se * rnorm(nsamp) #simulate effects under the alternative assumption
betahat    = rnorm(nsamp,beta,betahat.se)
zscore     = betahat/betahat.se
pval       = pchisq(zscore^2,df=1,lower.tail=FALSE)
plot(betahat,-log(pval))
```

Here is the EE analysis.
```{r}
  library(ashr)
  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
```

And here is how we can perform the ET analysis,
```{r}
  ashET.res = ash(betahat, betahat.se,method="fdr", alpha=1)
```

Now we compare the EE vs ET models:
```{r}
  ashEE.res$loglik-ashET.res$loglik
```

Then the log likelihood ratio is loglikEE-loglikET=`r ashEE.res$loglik-ashET.res$loglik`. This highly negative
loglikelihood indicate that the data strongly favor
the ET model, which is expected because the data were generated under this model.
(One might be tempted to ask whether the log likelihood ratio is ``significant". We don't
know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use EE, if it is negative then use ET.)

The above illustrates these ideas on simulations from the ET model. For comparison,
we now provide results for simulations under the EE model.
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = rnorm(nsamp) #simulate effects under the original assumption
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
  ashET.res = ash(betahat, betahat.se, method="fdr",alpha=1)
  ashEE.res$loglik-ashET.res$loglik
```
So here the log likelihood ratio is positive, indicating that the EE model is preferred
(which is expected since the data were generated under that model).

>>>>

```{r generate_data}
set.seed(1234)
nsamp      = 1000
betahat.se = rgamma(nsamp,1,1)
beta       = betahat.se * rnorm(nsamp)
betahat    = rnorm(nsamp,beta,betahat.se)
zscore     = betahat/betahat.se
pval       = pchisq(zscore^2,df=1,lower.tail=FALSE)
plot(betahat,-log(pval),pch = 20)
```

Here is the EE analysis.

```{r analysis_ee}
library(ashr)
ashEE.res = ash(betahat,betahat.se,method = "fdr",alpha = 0)
```

And here is how we can perform the ET analysis.

```{r analysis_et}
ashET.res = ash(betahat,betahat.se,method = "fdr",alpha = 1)
```

Now we compare the EE vs ET models:
```{r compare_analyses}
print(ashEE.res$loglik - ashET.res$loglik)
```

Then the log likelihood ratio is loglikEE - loglikET =
`r ashEE.res$loglik - ashET.res$loglik`. This highly negative loglikelihood
indicates that the data strongly favor the ET model, which is expected
because the data were generated under this model. (One might be tempted to ask
whether the log likelihood ratio is "significant". We don't know how to address
this question, but suggest in practice it doesn't matter: if the loglikelihood
ratio is positive then use EE, if it is negative then use ET.)

The above illustrates these ideas on simulations from the ET model. For
comparison, we now provide results for simulations under the EE model.

```{r generate_data_from_ee_model}
set.seed(1234)
samp       = 1000
betahat.se = rgamma(nsamp,1,1)
beta       = rnorm(nsamp)
betahat    = rnorm(nsamp,beta,betahat.se)
zscore     = betahat/betahat.se
pval       = pchisq(zscore^2,df = 1,lower.tail = FALSE)
ashEE.res  = ash(betahat,betahat.se,method = "fdr",alpha = 0)
ashET.res  = ash(betahat,betahat.se,method = "fdr",alpha = 1)
print(ashEE.res$loglik - ashET.res$loglik)
```

So here the log likelihood ratio is positive, indicating that the EE model
is preferred (which is expected since the data were generated under that
model).
